---
title: "PML Project"
author: "Jonathan Hélie"
date: "15/11/2020"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Library and seed

```{r}
library(caret)
library(rattle)
set.seed(1234567890)
```

## Loading data

The training data for this project are available here:
https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

The test data (used for the quiz) are available here:
https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

```{r}
data <- read.csv("D:/Documents/R Work Files/PML Week 4 project/pml-training.csv")
quiz <- read.csv("D:/Documents/R Work Files/PML Week 4 project/pml-testing.csv")
```

## Data slicing

Since the provided "testing" dataset doesn't include the outcome and is the subject of the quiz, I split the "training" dataset provided into 2 datasets, 1 for model training and the other one for testing/validation using the createDataPartition function.

```{r}
split <- createDataPartition(data$classe, times=1, p=0.7, list=FALSE)
testing <- data[-split,]
training <- data[split,]
```

## Data cleaning

I have apply the following treatment to the data in order to keep only the most interesting variables :

1. Remove the first 7 variables, because they are not movement related variables. They define the candidat, timestamp and more.
2. Remove columns with a lot of NA (90% and more)
3. Remove columns with a lot of missing values (90% and more)

The results of those treatment is a dataset containing 52 variables and the outcome.

```{r}
training <- training[,-7:-1]
training <- training[,colMeans(is.na(training))<0.9]
training <- training[,colMeans(training=="")<0.9]
```

## Model training and testing

In order to reduce the run time, I decided to change the Train function resampling method to K-fold with K equal to 5 parameters.

```{r}
trControl <- trainControl(method="cv", number=5)
```

I will try 3 type of models that could be appropriate for such a problem, where we have to predict a class, which are classification tree, random forest and gradient boosting.

For each model, I will :
1. Train it with the Train function, using the resampling method discuss previously and the training dataset
2. Predict the testing dataset and assess the accuracy

### Classification tree

```{r}
model1 <- train(classe ~ ., method = "rpart", data=training, trControl=trControl)
model1pred <- predict(model1,newdata=testing)
confMat1 <- confusionMatrix(testing$classe,model1pred)
confMat1$overall[1]
```

With an accuracy of 49.7%, this model is clearly bad. Such a classification tree model is probably to simple for a complex problem like that with a lot of possible variables.


### Random forest

```{r}
model2 <- train(classe ~ ., method = "rf", data=training, trControl=trControl)
model2pred <- predict(model2,newdata=testing)
confMat2 <- confusionMatrix(testing$classe,model2pred)
confMat2$overall[1]
```

This model perform very well with an accuracy of 99.4% over the testing dataset.


### Gradient boosting

```{r}
model3 <- train(classe ~ ., method = "gbm", data=training, trControl=trControl, verbose = FALSE)
model3pred <- predict(model3,newdata=testing)
confMat3 <- confusionMatrix(testing$classe,model3pred)
confMat3$overall[1]
```

This model perform quite well with an accuracy of 96.0%, but is not as good as the random forest one.


### Model selection

With an accuracy of 99.4%, the random forest perform very well and is the selected model to go through the quiz.


## Quiz prediction

```{r}
data.frame(number=quiz$X,answer=predict(model2,newdata=quiz))
```